Generative Software Development\\{\em\normalsize Research Offering

### 2023-01-11 Generative Software/proposal.tex
WIP

\section{Research question}
\label{sct:Research question}
   Data driven organizations that change frequently are facing the dilemma of buy vs. build.
   On one hand, flexible and fast adaptation requires immediate deployment of new software components, without interrupting the ongoing business.
   On the other hand, one-of-a-kind organizations require one-of-a-kind components that need to be developed first.
   Commercial off-the-shelf (COTS) components provide speedy deployments, but can be too generic.
   Homebrew components can satisfy specific requirements, but take time to build and maintain.
   The so called ``low code platforms'' are sometimes advocated as the solution
   because they generate user functionality from a graphical coding environment.
   However, practical experience shows that these platforms are not so ``low code'' after all,
   when the database structure is not properly aligned with the user functionality.

   In practice, we see many organizations that opt for homebrew solutions.
   Generative development can shorten the development time of homebrew software, regaining some of the desired flexibility.

\section{Plan}
   We want to participate in a larger research project by means of one researcher (PhD candidate, full-time, 4 years) and one scientific programmer.
   We intend to further the knowledge about generative software development by participating in a larger research effort.
   We bring the expertise~\cite{JoostenRAMiCS2017,Steenvoorden2022} and tools~\cite{Joosten-JLAMP2018,10.1145/3354166.3354182} on generative software development
   as an asset to that project at no extra cost.
   The authors also intend to participate in person in the role of supervisors of the researcher and the scientific programmer, each for 0.1 FTE.

>>>>

\section{Research question}
\label{sct:Research question}
   Data driven organizations that change frequently are facing the dilemma of buy vs. build.
   On one hand, flexible and fast adaptation requires immediate deployment of new software components, without interrupting the ongoing business.
   Commercial off-the-shelf (COTS) components can solve that problem, but they can be too generic.
   On the other hand, one-of-a-kind organizations require one-of-a-kind components that are not readily available.
   The so called ``low code an no code platforms'' are sometimes advocated as the solution
   because their (re-)sellers claim that the business itself can configure such systems.
   However, we have seen some disappointing results of that in practice, because of integration problems and maintainability issues.
   So homebrew solutions are still widespread in organizations with one-of-a-kind requirements.
   They are still willing to invest the time to build and maintain information systems and take the risk of a failing IT-innovation.

   Generative development can shorten the development time and reduce the development risks of homebrew software,
   regaining the desired flexibility and satisfying specific one-of-a-kind requirements at the same time.
   Current research has shown that complete working information systems can be generated.
   To get this technology ready for real use, research needs to be done on gaining more control over non-functional properties of generated systems,
   such as scalability, robustness, and security.

\section{Plan}
   We want to participate in a larger research project by means of one researcher (PhD candidate, full-time, 4 years) and one scientific programmer (professional, full-time, 4 years).
   We intend to further the knowledge about generative software development by participating in a larger research effort.
   We bring the expertise~\cite{JoostenRAMiCS2017,Steenvoorden2022} and tools~\cite{Joosten-JLAMP2018,10.1145/3354166.3354182} on generative software development
   as an asset to that project at no extra cost.
   The authors also intend to participate in person in the role of supervisors of the researcher and the scientific programmer, each for 0.1 FTE.
